---
title: "github_document"
author: "mm5951"
date: "2022-09-18"
output: github_document
---
#Load library
```{r, echo = FALSE}
library(tidyverse)
```

Welcome to mm5951 homework 1 solution!

#Load dataset
First, I'll load the provided dataset for this exercise.

```{r}
data("penguins", package = "palmerpenguins")
```

#Dataset description
Next, I will briefly describe the key characteristics of the "penguins" dataset. I'll first use the `skmr::skim`command to print out a summary of the dataset. The "penguins" dataset contains 344 rows and 8 columns. That corresponds to the data of n=8 different variables (columns) from n=344 penguins (rows).

The "penguins" dataset contains 5 numerical variables ("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g" and "year") and 3 categorical ones ("species", "island" and "sex"). 

As a summary, using the mean output, the average penguin will measure 43.92 mm in height and 17.15 mm in depth, have a flipper length of 200.92 mm and weight 4.2 kg. The overall dataset contains n=168 male and n=165 female from 3 main species (Adelie n=152, Gentoo n=124 and Chinstrap n=68). The data was collected between the years 2007-09.

```{r}
skimr::skim(penguins)
```

Alternatively, I could have used the `nrow`and `ncol` functions to return the number of rows and columns of the specified matrix ("penguins"), respectively.

```{r}
nrow(penguins)
ncol(penguins)
```


#Graphic outputs
