---
title: "Homework 1"
author: "mm5951"
date: "2022-09-18"
output: github_document
---
```{r, include = FALSE}
library(tidyverse)
```

Welcome to mm5951 homework 1 solution!

# Problem 1

## Load dataset

First, I'll **load the provided dataset** for this exercise.

```{r}
data("penguins", package = "palmerpenguins")
```

## Dataset description

Next, I will briefly describe the **key characteristics** of the **"penguins" dataset**. I'll first use the `skmr::skim`command to print out a summary of the dataset. The "penguins" dataset contains 344 rows and 8 columns. That corresponds to the data of n=8 different variables (columns) from n=344 penguins (rows).

The "penguins" dataset contains:
* 5 numerical variables ("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g" and "year"), and 
* 3 categorical ones ("species", "island" and "sex"). 

As a summary, using the mean output, the average penguin will measure 43.92 mm in height and 17.15 mm in depth, have a flipper length of 200.92 mm and weight 4.2 kg. The overall dataset contains n=168 male and n=165 female from 3 main species (Adelie n=152, Gentoo n=124 and Chinstrap n=68). The data was collected between the years 2007-09.

```{r}
skimr::skim(penguins)
```

Alternatively, I could have used the `nrow`and `ncol` functions to return the number of rows and columns of the specified matrix ("penguins"), respectively.

```{r}
nrow(penguins)
ncol(penguins)
```

## Graphic outputs

Next, with the `ggplot`function, I will **produce a scatter plot** of the variables flipper_length_mm (y) vs bill_length_mm (x), using different color points for each species.  

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

Finally, using the `ggsave`function I will **save the scatter plot** in .png format into the project directory.

```{r}
ggsave("scatterplot.png")
```

# Problem 2

## Create a data frame

First, I will create a data frame using the `tibble`function, incorporating the following elements:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r}
Hw1_df = tibble(
  norm = rnorm(n = 10),
  vec_logical = norm > 0,
  vec_char = c("This", "is", "homework", "one", "solution", "for", "data", "science", "class", "fall" ),
  vec_factor = factor(c("male", "male", "female", "female", "nonbinary", "nonbinary", "male", "female", "nonbinary", "male"))
)
```

## Calculate means

Next, I will calculate the mean of some of the objects as created above using the `mean`function. Note that the mean cannot be estimated for categorical variables.

Using the `pull`function, I will 

## Convert variable types

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?



