---
title: "Homework 1"
author: "mm5951"
date: "2022-09-18"
output: github_document
---
```{r, include = FALSE}
library(tidyverse)
```

Welcome to mm5951 homework 1 solution!

# Problem 1

## Load dataset

First, I'll **load the provided dataset** for this exercise.

```{r}
data("penguins", package = "palmerpenguins")
```

## Dataset description

Next, I will briefly describe the **key characteristics** of the **"penguins" dataset**. 

The "penguins" dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. That corresponds to the data of n=8 different variables (columns) from n=344 penguins (rows). As per homework request, the mean flipper lenght is `r round(mean(na.omit(penguins$flipper_length_mm)),2)` mm.

In order to obtain a comprehensive description of the dataset, I will use the `skmr::skim`function to print out a summary of the dataset (see chunk below, results not displayed, rather summarised). 

```{r, eval = FALSE}
skimr::skim(penguins)
```

The "penguins" dataset contains:
* 5 numerical variables ("bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g" and "year"), and 
* 3 factorial ones ("species", "island" and "sex"). 

As a summary, using the mean output, the average penguin will measure 43.92 mm in height and 17.15 mm in depth, have a flipper length of 200.92 mm and weight 4.2 kg. The overall dataset contains n=168 male and n=165 female from 3 main species (Adelie n=152, Gentoo n=124 and Chinstrap n=68). The data was collected between the years 2007-09.

## Graphic outputs

Next, with the `ggplot`function, I will **produce a scatter plot** of the variables flipper_length_mm (y) vs bill_length_mm (x), using different color points for each species.  

```{r, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
```

Finally, using the `ggsave`function I will **save the scatter plot** in .png format into the project directory.

```{r, message=FALSE}
ggsave("scatterplot.png")
```

# Problem 2

## Create a data frame

First, I will create a data frame using the `tibble`function, incorporating the following elements:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r}
Hw1_df = tibble(
  samp = rnorm(n = 10),
  vec_logical = samp > 0,
  vec_char = c("This", "is", "homework", "one", "solution", "for", "data", "science", "class", "fall" ),
  vec_factor = factor(c("male", "male", "female", "female", "nonbinary", "nonbinary", "male", "female", "nonbinary", "male"))
)
```

## Calculate means

Next, I will calculate the mean of some of the objects as created above using the `mean` function. To do so, first I will pull each variable (using the function `pull`)and name them y1-y4 following the order in which they appear in the dataframe.

```{r}
y1 = pull(.data = Hw1_df, var = samp)
y2 = pull(.data = Hw1_df, var = vec_logical)
y3 = pull(.data = Hw1_df, var = vec_char)
y4 = pull(.data = Hw1_df, var = vec_factor)
```

Then the means are calculated. For the numerical variable, results will be rounded to 3 decimals using the `round` function.

```{r}
round(
  mean(y1), 3
  )

mean(y2)
mean(y3)
mean(y4)
```

As a result of the chunk code, I obtain the mean for "y1 = samp" (data type: numerical) and "y2 = vec_logical" (data type: logical), which are 0.069 and 0.5, respectively. Note that given that y1 is a random sample, would the session be re-started and the .Rmd re-knitted, values might change.

Conversely, I get a warning message for the remaining two variables, given that the data type of some of the vectors is not numerical nor logical, rather character ("y3 = vec_char") or factor ("y4 = vec_factor").

## Convert variable types

Finally, I will convert the logical, character, and factor variables into numerical using the `as.numeric` function. For each variable, I will use the `pull`function, so that they can be addressed separately from the "Hw1_df" dataframe. Note the chunk is shown, but not the output, as per homework instructions. 

```{r, eval = FALSE}
as.numeric(y2)
as.numeric(y3)
as.numeric(y4)
```

The "vec_logical" and "vec_factor" are coerced into numeric elements, following a logical rationale (for instance, in "vec_logical", TRUE=1 and FALSE=0; in "vec_factor", its 3 levels are converted into female=1, male=2, nonbinary=3 according to their alphabetical order). The character vector "vec_char" is more challenging, given that there's no logical conversion possible with the provided criteria, and R coerces its 10 elements to NAs.

The kind of data contained in each variable determines whether a mean can be calculated. In general, means can be calculated for numerical and logical vectors (in the latter, using the TRUE=1 and FALSE=0 coercion). Once converted into numerical, the mean of the remaining two vectors could be calculated (not required as per homework instructions).

That's it for homework 1!

